---
title: "Urban Pedals: Unveiling Toronto's Bike Share Story"
output:
  html_document:
    df_print: paged
  html_notebook: default
  pdf_document: default
---

## Overview
This report contains the code and data for a data analysis project focused on exploring the ridership patterns of the bike-share scheme in Toronto. The analysis is conducted using the R programming language to gain insights into user behaviors, popular routes, and temporal trends.

---

## Data Source

- Ref:
  - https://open.toronto.ca/dataset/bike-share-toronto-ridership-data/

---

## Download and Import data.

- Import libraries.


```{r}
# load library
library(opendatatoronto)
library(dplyr)
library(tidyr)
library(ggplot2)
```

- Manually download data files from data source.
  - We have 52 files of 5-year data, from 2019 to 2023.
  - List of all data files
  
```{r}
# Define path of dataset
DATA_PATH <- "./source/"

# get the paths for each data file.
FILE_PATH_LIST <- list.files(DATA_PATH, pattern = "\\.csv$", full.names = TRUE)
FILE_PATH_LIST
```

- Import data from data files.

  - Check the columns of each csv file for data consisitence.
  - Combine all data into a uniformed dataframe for further anaysis.


```{r}
# # Apply read.csv, a function to import data from csv file, for each file.
# # Get a list of df
# df_list <- lapply(FILE_PATH_LIST, read.csv)
# 
# 
# # Union all df by row
# raw_df <- do.call(rbind, df_list)
# raw_df
```

---

- Test using selective data

```{r}
# paths <- c(
#   "./source/2019-Q1.csv",
#   "./source/2019-Q2.csv",
#   "./source/2019-Q3.csv",
#   "./source/2019-Q4.csv",
#   "./source/2020-01.csv",
#   "./source/2020-02.csv",
#   "./source/2020-03.csv",
#   "./source/2020-04.csv",
#   "./source/2020-05.csv",
#   "./source/2020-06.csv",
#   "./source/2020-07.csv",
#   "./source/2020-08.csv",
#   "./source/2020-09.csv",
#   "./source/2020-10.csv",
#   "./source/2020-11.csv",
#   "./source/2020-12.csv"
# )
# # paths <- c("./source/2020-01.csv")
# df_list <- lapply(paths, read.csv)
# raw_df <- do.call(rbind, df_list)
# raw_df
```

---

## Data Processing

### Handling NA and NULL value

- Handle NA value that exist in the raw data

```{r}
# # Remove rows with any NA values
# proc_NA_df <- raw_df[complete.cases(raw_df), ]
# 
# # Remove rows with "NULL" values
# proc_null_df <- proc_NA_df %>%
#   filter(!(End.Station.Name == "NULL" | End.Station.Id == "NULL" | Start.Station.Id =="NULL" | Start.Station.Name == "NULL"))
# proc_df <- proc_null_df
```

---

### Converting Time-Dimension Data 

- Divide time into year, month, date, hour, and minute.

```{r}
# # Divide "Start.Time" into columns
# proc_df$Start.Time <- as.POSIXct(proc_df$Start.Time, format = "%m/%d/%Y %H:%M")
# proc_df$Start.Year <- as.factor(format(proc_df$Start.Time, "%Y"))
# proc_df$Start.Month <- as.factor(format(proc_df$Start.Time, "%m"))
# proc_df$Start.Date <- as.factor(format(proc_df$Start.Time, "%d"))
# proc_df$Start.Hours <- as.factor(format(proc_df$Start.Time, "%H"))
# proc_df$Start.Minutes <- as.factor(format(proc_df$Start.Time, "%M"))
# 
# # Divide "End" into columns
# proc_df$End.Time <- as.POSIXct(proc_df$End.Time, format = "%m/%d/%Y %H:%M")
# proc_df$End.Year <- as.factor(format(proc_df$End.Time, "%Y"))
# proc_df$End.Month <- as.factor(format(proc_df$End.Time, "%m"))
# proc_df$End.Date <- as.factor(format(proc_df$End.Time, "%d"))
# proc_df$End.Hours <- as.factor(format(proc_df$End.Time, "%H"))
# proc_df$End.Minutes <- as.factor(format(proc_df$End.Time, "%M"))
# 
# # factor user.type
# proc_df$User.Type <- as.factor(proc_df$User.Type)
# 
# 
# # Drop Start.Time and End.Time
# proc_df <- proc_df %>% select(-Start.Time, -End.Time)

```

- Check NA value again, in case of any possible values generated during the data processing.

```{r}
# is_miss <- any(is.na(proc_df))
# 
# # if the processed_df contains missing value, drop the rows with missing values and assign to df
# if (is_miss) {
#   df <- proc_df[complete.cases(proc_df), ]
# # otherwise, df = proc_df
# }else{
#   df <- proc_df
# }
# 
# is_miss <- any(is.na(df))
# cat("df has missing value? ", is_miss)  # output result
  
```

- Data Overview after data processing.

```{r}
# # Data overview after data processing
# num_row <- nrow(df)             # total rows
# column_names <- colnames(df)    # column names
# cat("\n\nNumber of rows: ", "\n", num_row)
# cat("\n\nColumn names: ", "\n", column_names)
# 
# cat("\n\nDisplay the Structure:\n")
# str(df)
# 
# cat("\n\nDisplay Summaries:\n")
# summary(df)
```

- Preview data

```{r}
# df
# head(df, 10)
```

---

### Exporting Processed Data(Optional)

- Export the Processed data to permanently store the processed data.

```{r}
# # Path to export
# output_file <- "./data/dataset.csv"
# # export
# write.csv(df, file = output_file, row.names = FALSE)
```

---

### ReLoading Processed Data

```{r}
data_file <- "./data/dataset.csv"
df <- read.csv(data_file)
df
```


- Data Overview afater loading data.

```{r}
# Data overview after data processing
num_row <- nrow(df)             # total rows
column_names <- colnames(df)    # column names
cat("\n\nNumber of rows: ", "\n", num_row)
cat("\n\nColumn names: ", "\n", column_names)

cat("\n\nDisplay the Structure:\n")
str(df)

cat("\n\nDisplay Summaries:\n")
summary(df)
```

- Preview data

```{r}
# df
head(df, 10)
```

## Trip analysis

### Yearly Trip Trends

```{r}
trip_yearly <- df %>%
  group_by(Year = Start.Year) %>%
  summarize(Trip = n())
trip_yearly
```

```{r}
ggplot(
  data = trip_yearly, 
  mapping = aes(
      x = Year, 
      y = Trip,
      group = 1
  )
) +
  geom_line() + 
  geom_point() +
  geom_text(aes(label = as.character(Trip)), vjust = 1, hjust = 0.5) +
  labs(
    title = "Yearly Trip Trend",
    x = "Year",
    y = "Trip"
  ) +
  theme_minimal()

```

---

### Monthly Trip Distribution

- Comparing monthly trip to unveil patterns over the months

```{r}
trip_monthly_across_year <- df %>%
  group_by(
    Year = Start.Year, 
    Month = Start.Month
    ) %>%
  summarize(Trip = n())
trip_monthly_across_year
```


```{r}
ggplot(
  data = trip_monthly_across_year, 
  mapping = aes(
    x = factor(Month), 
    y = Trip, 
    fill = factor(Month)
    )
  ) +
  geom_boxplot(show.legend = FALSE) +
  labs(
    title = "Monthly Distribution of Trips",
    x = "Month",
    y = "Trip",
    fill = "Month"
  ) +
  theme_minimal()
```

---

### Hourly Trip Pattern

- Exploring Patterns Throughout the Day


```{r}
trip_hourly_across_month <- df %>%
  group_by(
    Month = Start.Month, 
    Hour = Start.Hours
    ) %>%
  summarize(Trip = n())
trip_hourly_across_month
```


```{r}
ggplot(
  trip_hourly_across_month, 
  aes(
    x = factor(Hour), 
    y = Trip, 
    fill = factor(Hour)
    )
  ) +
  geom_boxplot(show.legend = FALSE) +
  labs(
    title = "Hourly Trips Distribution",
    x = "Hour of Day",
    y = "Number of Trips",
    fill = "Hour of Day") +
  theme_minimal()
```

---

## Duration Analysis

### Duration Yearly Trend

```{r}
duration_yearly <- df %>%
  group_by(Year = Start.Year) %>%
  summarize(Mean_Duration = round(mean(Trip..Duration) / 60, 2))  

duration_yearly

```


```{r}
ggplot(
  data = duration_yearly, 
  mapping = aes(
      x = Year, 
      y = Mean_Duration,
      group = 1
    )
  ) +
  geom_line() + 
  geom_point() +
  geom_text(
    aes(
      label = as.character(Mean_Duration)
      ), 
    vjust = 1, 
    hjust = 0.5
  ) +
  labs(
    title = "Average Duration Trend",
    x = "Year",
    y = "Average Duration"
  ) +
  theme_minimal()
```

---

### Duration Pattern over Months

```{r}
duration_monthly_across_year <- df %>%
  group_by(
    Year = Start.Year, 
    Month = Start.Month) %>%
  summarize(Mean_Duration = round(mean(Trip..Duration) / 60, 2))  # Convert mean to minutes

duration_monthly_across_year
```

```{r}
ggplot(
  data = duration_monthly_across_year, 
  mapping = aes(
    x = factor(Month), 
    y = Mean_Duration, 
    fill = factor(Month)
    )
  ) +
  geom_boxplot(show.legend = FALSE) +
  labs(
    title = "Monthly Duration Distribution",
    x = "Month",
    y = "Mean Duration (minutes)"
  ) +
  theme_minimal()
```

---

### Duration Pattern throughout A Day

```{r}
duration_hourly_across_month <- df %>%
  group_by(
    Month = Start.Month, 
    Hour = Start.Hours
    ) %>%
  summarize(Mean_Duration = round(mean(Trip..Duration) / 60, 2))  # Convert mean to minutes

duration_hourly_across_month
```

```{r}
ggplot(
  data = duration_hourly_across_month, 
  mapping = aes(
    x = factor(Hour), 
    y = Mean_Duration, 
    fill = factor(Hour)
    )
  ) +
  geom_boxplot(show.legend = FALSE) +
  labs(
    title = "Hourly Duration Distribution",
    x = "Hour of Day",
    y = "Mean Duration (minutes)"
  ) +
  theme_minimal()
```

---

## Geolocation Analysis

### Hot Start Locations

#### Hot start spot (Top 10)
```{r}

hot_start_spot <- df %>%
  group_by(Start.Station.Name) %>%
  summarize(Total_Trips = n()) %>% 
  arrange(desc(Total_Trips)) %>% 
  slice_head(n = 10)
hot_start_spot

```
```{r}
ggplot(
  data = hot_start_spot, 
  mapping = aes(
      x = reorder(Start.Station.Name, -Total_Trips), 
      y = Total_Trips, 
      fill = reorder(Start.Station.Name, -Total_Trips)
      )
  ) +
  geom_bar(
    stat = "identity"
  ) +
  labs(
    title = "Total Number of trips per starting stations",
    x = "Starting Stations",
    y = "Number of trips",
    fill = "Starting Station"
  ) +
  theme_minimal()+
  theme(axis.text.x = element_blank()) 


```

#### Hot start street (Top 10)

```{r}

library(stringr)

hot_start_streets <- df %>%
  separate_rows(Start.Station.Name, sep = " / ") %>% # Separate the station names 
  mutate(Start.Station.Street = Start.Station.Name) %>%
  group_by(Start.Station.Street) %>%
  summarize(Total_Trips = n()) %>%
  arrange(desc(Total_Trips)) %>% 
  slice_head(n = 10)

hot_start_streets

```
```{r}
ggplot(
  data = hot_start_streets, 
  mapping = aes(
      x = reorder(Start.Station.Street, -Total_Trips), 
      y = Total_Trips, 
      fill = reorder(Start.Station.Street, -Total_Trips)
      )
  ) +
  geom_bar(
    stat = "identity"
  ) +
  labs(
    title = "Total Number of trips per starting street",
    x = "Starting street",
    y = "Number of trips",
    fill = "Starting street"
  ) +
  theme_minimal()+
  theme(axis.text.x = element_blank()) 
```

### Hot End Locations

#### Hot End spot (Top 10)

```{r}

hot_end_spot <- df %>%
  group_by(End.Station.Name) %>%
  summarize(Total_Trips = n()) %>% 
  arrange(desc(Total_Trips)) %>% 
  slice_head(n = 10)
hot_end_spot

```
```{r}
ggplot(
  data = hot_end_spot, 
  mapping = aes(
      x = reorder(End.Station.Name, -Total_Trips), 
      y = Total_Trips, 
      fill = reorder(End.Station.Name, -Total_Trips)
      )
  ) +
  geom_bar(
    stat = "identity"
  ) +
  labs(
    title = "Total Number of trips per ending station",
    x = "Ending station",
    y = "Number of trips",
    fill = "Ending station"
  ) +
  theme_minimal()+
  theme(axis.text.x = element_blank()) 
```

#### Hot end street (Top 10)

```{r}

hot_end_streets <- df %>%
  separate_rows(End.Station.Name, sep = " / ") %>% # Separate the station names 
  mutate(End.Station.Street = End.Station.Name) %>%
  group_by(End.Station.Street) %>%
  summarize(Total_Trips = n()) %>%
  arrange(desc(Total_Trips)) %>% 
  slice_head(n = 10)

hot_end_streets


```
```{r}
ggplot(
  data = hot_end_streets, 
  mapping = aes(
      x = reorder(End.Station.Street, -Total_Trips), 
      y = Total_Trips, 
      fill = reorder(End.Station.Street, -Total_Trips)
      )
  ) +
  geom_bar(
    stat = "identity"
  ) +
  labs(
    title = "Total Number of trips per ending street",
    x = "Ending street",
    y = "Number of trips",
    fill = "Ending street"
  ) +
  theme_minimal()+
  theme(axis.text.x = element_blank()) 
```

### Hot Trips (Top 10)

```{r}


hot_trip <- df %>%
  mutate(Trip.Spots = paste(Start.Station.Name, End.Station.Name, sep =" -> ")) %>%
  group_by(Trip.Spots) %>%
  summarize(Total_Trips = n()) %>% 
  arrange(desc(Total_Trips)) %>% 
  slice_head(n = 10)
hot_trip 


```
```{r}
ggplot(
  data = hot_trip, 
  mapping = aes(
      x = reorder(Trip.Spots, -Total_Trips), 
      y = Total_Trips, 
      fill = reorder(Trip.Spots, -Total_Trips)
      )
  ) +
  geom_bar(
    stat = "identity"
  ) +
  labs(
    title = "Total Number of trips per origin-destination pair",
    x = "Origin-destination pair",
    y = "Number of trips",
    fill = "Origin-destination pair"
  ) +
  theme_minimal()+
  theme(axis.text.x = element_blank()  ,
        legend.text = element_text(size = 6))
```


---

## Shiny Rendering

```{r}
library(shiny)
library(ggplot2)

ui <- fluidPage(
  tags$head(
    tags$style(HTML("
      #sidebar {
        position: fixed;
        top: 0;
        left: 0;
        bottom: 0;
        width: 300px; /* Set your desired width */
        background-color: #f8f9fa; /* Set your desired background color */
        padding: 10px;
        
      },
      #mainPanel {
        font-size: 24px;
        padding-right: 200px;
      }
    "))
  ),
  
  sidebarLayout(
    sidebarPanel(
      id = "sidebar",
      tags$div(
        tags$h3(tags$a(href = "#title","Urban Pedals")),
        tags$ul(
          tags$li(
            tags$a(href = "#source", "Data Source")
          ),
          tags$li(
            tags$a(href = "#processing", "Data Processing")
          ),
          # Trip
          tags$li(
            tags$a(href = "#trip", "Trip Analysis"),
            tags$ul(
              tags$li(tags$a(href = "#yearly-trip", "Yearly Trends")),
              tags$li(tags$a(href = "#monthly-trip", "The Pattern over Months")),
              tags$li(tags$a(href = "#hourly-trip", "The Pattern throughout A Day"))
            )
          ),
          # duration
          tags$li(tags$a(href = "#duration", "Duration Analysis"),
                  tags$ul(
                    tags$li(tags$a(href = "#yearly-duration", "Yearly Trend")),
                    tags$li(tags$a(href = "#monthly-duration", "The Pattern over Months")),
                    tags$li(tags$a(href = "#hourly-duration", "The Pattern throughout A Day"))
                  )
          ),
          # location
          tags$li(tags$a(href = "#location", "Location Analysis"),
                  tags$ul(
                    tags$li(tags$a(href = "#start-street", "Count of start street")),
                    tags$li(tags$a(href = "#end-street", "Count of end street")),
                    tags$li(tags$a(href = "#trip-pairs", "Count of trip pairs"))
                  )
          )
        )
      )
    ),
    mainPanel(
      style = "padding-right: 100px; text-align: left;",
      
      h1("Urban Pedals:", align = "center", id = "title"),
      h3("Unveiling Toronto’s Bike Share Story", align = "center"),
      h2("Data Source", id = "source"),
      div(
        style = "font-size: 18px",
        p("Historical Bike Share Toronto ridership data."),
        p("Ref: ",tags$a(href ="https://open.toronto.ca/dataset/bike-share-toronto-ridership-data/", "Toronto Open Data")
        ),
      ),
      tags$hr(),
      div(style = "height: 50px"),

      # data processing
      h2("Data Processing", id = "processing"),
      div(
        style = "font-size: 18px",
        p("Handle NA and 'NULL' Value"),
        p("Data Overview:"),
      ),
      br(),
      imageOutput("photo"),
      tags$hr(),


      # Trip Analysis
      h2("Trip analysis", id = "trip"),
      h3("Yearly Trends", id = "yearly-trip"),
      tableOutput('trip_yearly_tb'),
      br(),
      plotOutput("trip_yearly_plot"),
      br(),
      div(
        style = "font-size: 24px;",
        p("The annual trip count exhibits a consistent upward trend, doubling from approximately 2.4 million to 4.6 million over the specified period.")
        ),
      tags$hr(),
      
      # trip monlty
      h3("The Pattern over Months", id = "monthly-trip"),
      br(),
      DT::dataTableOutput("trip_monthly_tb"),
      br(),
      plotOutput("trip_monthly_plot"),
      div(
        style = "font-size: 24px;",
        # description
        p("")
        ),
      tags$hr(style = "padding-bottom: 50px"),
      
      # trip hourly
      h3("The Pattern throughout A Day", id = "hourly-trip"),
      br(),
      DT::dataTableOutput("trip_hourly_tb"),
      br(),
      plotOutput("trip_hourly_plot"),
      div(
        style = "font-size: 24px;",
        # description
        p("")
        ),
      tags$hr(),
      
      # Duration
      h2("Duration Analysis", id = "duration"),
      
      # duration yearly
      h3("Yearly Trend", id = "yearly-duration"),
      br(),
      DT::dataTableOutput("duration_yearly_tb"),
      br(),
      plotOutput("duration_yearly_plot"),
      div(
        style = "font-size: 24px;",
        # description
        p("")
        ),
      tags$hr(),
      
      # duration monthly
      h3("The Pattern over Months", id = "monthly-duration"),
      br(),
      DT::dataTableOutput("duration_monthly_tb"),
      br(),
      plotOutput("duration_monthly_plot"),
      div(
        style = "font-size: 24px;",
        # description
        p("")
        ),
      tags$hr(),
      
      # duration hourly
      h3("The Pattern throughout A Day", id = "hourly-duration"),
      br(),
      DT::dataTableOutput("duration_hourly_tb"),
      br(),
      plotOutput("duration_hourly_plot"),
      div(
        style = "font-size: 24px;",
        # description
        p("")
        ),
      tags$hr(),
            
      
      # Location
      h2("Location Analysis", id = "location"),
      
      # location start street 
      h3("Total Number of trips per starting stations", id = "start-street"),
      br(),
      DT::dataTableOutput("hot_start_streets_tb"),
      br(),
      plotOutput("hot_start_streets_plot"),
      div(
        style = "font-size: 24px;",
        # description
        p("")
        ),
      tags$hr(),
            
      # location end street
      h3("Total Number of trips per ending street", id = "end-street"),
      br(),
      DT::dataTableOutput("hot_end_streets_tb"),
      br(),
      plotOutput("hot_end_streets_plot"),
      div(
        style = "font-size: 24px;",
        # description
        p("")
        ),
      tags$hr(),
            
      # location pairs
      h3("Total Number of trips per origin-destination pair", id = "trip-pairs"),
      br(),
      DT::dataTableOutput("hot_trip_tb"),
      br(),
      plotOutput("hot_trip_plot"),
      div(
        style = "font-size: 24px;",
        # description
        p("")
        ),
      tags$hr(),
      
      
    )
  )
)

server <- function(input, output) {
  
  output$photo <- renderImage({
    list(
      src = file.path("info.png"),
      contentType = "image/png",
      width = 800
      )
  }, deleteFile = FALSE)
  
  # trip yearly
  output$trip_yearly_tb <- renderTable(trip_yearly)
  output$trip_yearly_plot <- renderPlot({
    ggplot(
     data = trip_yearly,
     mapping = aes(
     x = Year,
     y = Trip,
     group = 1
     )
    ) +
     geom_line() +
     geom_point() +
     geom_text(aes(label = as.character(Trip)), vjust = 1, hjust = 0.5) +
     labs(
     title = "Yearly Trip Trend",
     x = "Year",
     y = "Trip"
     ) +
     theme_minimal()
  })
  
  # trip monthly
  output$trip_monthly_tb <- DT::renderDataTable(DT::datatable({
    data <- trip_monthly_across_year
    data
  }))
  
  output$trip_monthly_plot <- renderPlot({
    ggplot(
       data = trip_monthly_across_year,
       mapping = aes(
       x = factor(Month),
       y = Trip,
       fill = factor(Month)
       )
       ) +
       geom_boxplot(show.legend = FALSE) +
       labs(
       title = "Monthly Distribution of Trips",
       x = "Month",
       y = "Trip",
       fill = "Month"
       ) +
       theme_minimal()
  })

  # trip hourly
  output$trip_hourly_tb <- DT::renderDataTable(DT::datatable({
    data <- trip_hourly_across_month
    data
  }))
  
  output$trip_hourly_plot <- renderPlot({
    ggplot(
      trip_hourly_across_month,
      aes(
      x = factor(Hour),
      y = Trip,
      fill = factor(Hour)
      )
      ) +
      geom_boxplot(show.legend = FALSE) +
      labs(
      title = "Hourly Trip Distribution",
      x = "Hour of Day",
      y = "Number of Trips",
      fill = "Hour of Day") +
      theme_minimal()
  })
  
  # duration yearly
  output$duration_yearly_tb <- DT::renderDataTable(DT::datatable({
    data <- duration_yearly
    data
  }))
  
  output$duration_yearly_plot <- renderPlot({
    ggplot(
      data = duration_yearly,
      mapping = aes(
      x = Year,
      y = Mean_Duration,
      group = 1
      )
      ) +
      geom_line() +
      geom_point() +
      geom_text(
      aes(
      label = as.character(Mean_Duration)
      ),
      vjust = 1,
      hjust = 0.5
      ) +
      labs(
      title = "Average Duration Trend",
      x = "Year",
      y = "Average Duration"
      ) +
      theme_minimal()
  })
  
  
 # duration monthly
  output$duration_monthly_tb <- DT::renderDataTable(DT::datatable({
    data <- duration_monthly_across_year
    data
  }))
  
  output$duration_monthly_plot <- renderPlot({
    ggplot(
     data = duration_monthly_across_year,
     mapping = aes(
     x = factor(Month),
     y = Mean_Duration,
     fill = factor(Month)
     )
     ) +
     geom_boxplot(show.legend = FALSE) +
     labs(
     title = "Monthly Duration Distribution",
     x = "Month",
     y = "Mean Duration (minutes)"
     ) +
     theme_minimal()
  })
  
  
  # duration hourly
  output$duration_hourly_tb <- DT::renderDataTable(DT::datatable({
    data <- duration_hourly_across_month
    data
  }))
  
  output$duration_hourly_plot <- renderPlot({
    ggplot(
       data = duration_hourly_across_month,
       mapping = aes(
       x = factor(Hour),
       y = Mean_Duration,
       fill = factor(Hour)
       )
       ) +
       geom_boxplot(show.legend = FALSE) +
       labs(
       title = "Hourly Duration Distribution",
       x = "Hour of Day",
       y = "Mean Duration (minutes)"
       ) +
       theme_minimal()
  })

  # location start street
  output$hot_start_streets_tb <- DT::renderDataTable(DT::datatable({
    data <- hot_start_streets
    data
  }))
  
  output$hot_start_streets_plot <- renderPlot({
  ggplot(
    data = hot_start_streets, 
    mapping = aes(
        x = reorder(Start.Station.Street, -Total_Trips), 
        y = Total_Trips, 
        fill = reorder(Start.Station.Street, -Total_Trips)
        )
    ) +
    geom_bar(
      stat = "identity"
    ) +
    labs(
      title = "Total Number of trips per starting street",
      x = "Starting street",
      y = "Number of trips",
      fill = "Starting street"
    ) +
    theme_minimal()+
    theme(axis.text.x = element_blank()) 
  })
  
  # location end street
  output$hot_end_streets_tb <- DT::renderDataTable(DT::datatable({
    data <- hot_end_streets 
    data
  }))
  
  output$hot_end_streets_plot <- renderPlot({
  ggplot(
    data = hot_end_streets, 
    mapping = aes(
        x = reorder(End.Station.Street, -Total_Trips), 
        y = Total_Trips, 
        fill = reorder(End.Station.Street, -Total_Trips)
        )
    ) +
    geom_bar(
      stat = "identity"
    ) +
    labs(
      title = "Total Number of trips per ending street",
      x = "Ending street",
      y = "Number of trips",
      fill = "Ending street"
    ) +
    theme_minimal()+
    theme(axis.text.x = element_blank()) 
  })
  
  # location pairs
  output$hot_trip_tb <- DT::renderDataTable(DT::datatable({
    data <- hot_trip
    data
  }))
  
  output$hot_trip_plot <- renderPlot({
  ggplot(
    data = hot_trip, 
    mapping = aes(
        x = reorder(Trip.Spots, -Total_Trips), 
        y = Total_Trips, 
        fill = reorder(Trip.Spots, -Total_Trips)
        )
    ) +
    geom_bar(
      stat = "identity"
    ) +
    labs(
      title = "Total Number of trips per origin-destination pair",
      x = "Origin-destination pair",
      y = "Number of trips",
      fill = "Origin-destination pair"
    ) +
    theme_minimal()+
    theme(axis.text.x = element_blank()  ,
          legend.text = element_text(size = 6))
  })
  
  
}

shinyApp(ui, server)
```





